Informative Autonomous Sampling of Oceanographic Variables Using Joint Excursion Sets

### 2020-05-15 ans_to_reviewers/answers-to-reviewers.tex

I think this is an interesting paper. The paper is well written, 
and the authors have motivated the application well. 
The authors indicate several paths forward for work in this area, which I think pose interesting questions. They may want to elaborate a bit more on those.

\action: Space-time \cite{sigrist2015stochastic,richardson2017sparsity}. More complex sampling Look-ahead, going further. 3D sampling, requiring more computationally efficient spatial statistical models.

\item{\revcom - Minor 1:}\label{r3c2}

I think the equation $V_{m,upd}$ needs a j as a part of that subscript.

@@

\action: Our action

\item{\revcom - Major 2:}\label{r3c3}

Figure 4: Why not show excursion probabilities for, say, myopic selection as well?

### 2020-05-23 main_new.tex
Updates from Jo
Updated introduction

\address{\\Trygve Olav Fossum \\Department of Marine Technology\\ Otto Nielsens veg. 10, 7491 Trondheim\\ Norway\\
\printead{e1}}
\address{Jo Eidsvik\\Department of Mathematical Sciences\\ Hogskoleringen 1, 7491 Trondheim\\ Norway\\ \printead{e2}}
\address{David Ginsbourger\\ Idiap Research Institute\\ Rue Marconi 19, 1920 Martigny\\ Switzerland\\
\printead{e3}}
\address{Kanna Rajan\\Underwater Systems and Technology Laboratory,
  Faculty of Engineering,\\ Rua Dr. Roberto Frias\\ University of Porto, Portugal\\
% Department of Engineering Cybernetics\\ O. S. Bragstads Plass 2D, 7034 Trondheim\\ Norway\\
\printead{e4}}

@@

%Motivated by the challenges related to efficient allocation of sampling resources in environmental sensing, the combination of Excursion Probabilities and Gaussian process modeling is explored for autonomous robotic sampling of ocean features; enabling information driven measures  sampling efforts to high-interest regions. These regions are usually characterized by gradients of measurable environmental variables, e.g., temperature or salinity gradients, on which EPs subsequently can be used...Correlation among samples and multivariate requirements are typical in environmental studies.
\end{abstract}

%\begin{keyword}[class=MSC]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

@@

Monitoring the world's oceans has gained increased importance in 
light of the changing climate and increasing anthropogenic impact. A
central problem for understanding these factors is the lack of
representative data with sufficient resolution. Most of this
\emph{undersampling} can be attributed to the large spatio-temporal
variations on which ocean processes transpire, prompting the need for
effective means of sampling. By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use of new technologies including autonomous platforms can be combined with statistical methods to pursue measurements with high scientific relevance. In this paper we build upon the potential that lies in using autonomous underwater vehicles (AUVs) to conduct efficient sampling in the ocean \citep{das11b,Das2015,fossuminformation,fossum18b}. %Addressing this, the combination of statistical tools and robotic platforms is a natural symbiosis which enables information-based environmental sensing.

There are several obstacles when doing sampling with an AUV in the ocean, and we discuss some of these to motivate our setting, and also the modeling and methodological choices that will be made. First, full numerical ocean models based on complex differential equations cannot provide accurate results for online operations because the onboard computers lack the computational power required. Hence statistical proxy models of the environment must be used on the AUV. For our purpose, we rely on Gaussian process (GP) representations of the ocean variables of interest. Second, %with limited available information about the state of the ocean, 
with undersampling issues one must usually plan for active learning during the operation, where a relevant criterion is used to extract the most valuable designs. %, preferable in such a way that there is substantial value in reacting to information obtained from measurements taken in-situ. 
The acquired in-situ information must then be assimilated into the statistical model onboard the AUV, so that it can be used to re-compute the sampling criterion and in this way inform decisions on where to sample next. We use a criterion focusing on the prediction of excursion sets (ESs) for the multivariate spatial processes of interest. Third, the number of sampling designs is enormous, creating a trade-off between optimization and computability. One must often compromise in both development and practical operations, and we will rely on short-horizon optimization routines for sequential designs among nearest neighbors points in a spatial grid created for the AUV operation.

% After reviews: - This is moved to Section 2.1.

%From an oceanographic perspective, interesting regions are usually directly tied to a distinct phenomena that is of scientific interest. Each phenomenon can in turn be characterized by a set of process specific conditions expressed through different measures of key environmental variables, such as temperature or salinity. One such measure is the gradient, that can be associated with a number of important processes, such as the vertical location of the thermocline and pycnocline, location of upwelling systems, vertical mixing, eddies, fronts, and currents \cite{sverdrup2006}, as well as distribution, growth, and accumulation of biological activity \cite{SatOceanSoci00, Ryan2014}. These gradients create boundaries separating the ocean into process specific regions which are of profound interest to both identify and map effectively. Quantification of gradient features is therefore a much needed competence in robotic sampling.

%The imperfect understanding of causality coupled with the underlying fundamental uncertainty of the ocean implies that we need to obtain more measurements and use statistical models and methods to determine what and how phenomena form, to increase the quality of predictive statements.

We design AUV sampling strategies for reducing the spatially-integrated
Bernoulli variance (IBV) in the ES indicator variables. 
Recent work on ESs and associated excursion probabilities (EPs) connected to spatial statistics include
\cite{french2013spatio,bolin2015excursion,french2016credible}.
Our focus is on uncertainty
characterization and how it is influenced by sampling. In this sense,
our work is similar to \cite{bect2012,chevalier2014fast,azzimonti2016quantifying} who describe analytical results for
the IBV for univariate processes. A major contribution of our work is
to derive closed-form results for this design criteria for situations
where the underlying model is based on multivariate GPs. Another contribution is that we embed this approach in a sequential strategy to produce algorithms, and in applying this on an AUV in the real-world
scenario of autonomously sampling temperature and salinity variables in a river plume.

Motivating examples relevant for ESs of multivariate processes are abundant. In medicine, doctors do not rely solely on a single symptom but must see several combined effects before making a diagnosis. In our context of environmental applications, an example is that of photosynthesis where plants and algae produce oxygen only when exposed to enough light, carbon dioxide and water. Our key motivation in this paper is that temperature and salinity, combined, drive several processes in the oceans. For instance, salmon lice larvae are known to depend on relatively high temperature and high salinities to survive. Salmon lice are parasites to salmon, and the bivariate ES for temperature and salinity is hence important to understand for sustainable salmon fisheries management. The salinity and temperature excursions of a river plume can similarly help characterize its oceanographic and bio-geochemical processes \cite{hopkins2013detection,Pinto2018}, and sampling should be aimed at mapping relevant parts to reduce uncertainty in the temperature and salinity ES. Unlike what has been done in previous plume exploration studies, we emphasize the links to spatial statistical modeling and targeted multivaraite sampling criteria for sequential sampling. Addressing this, the combination of statistical tools and robotic platforms is a natural symbiosis which enables information-based environmental sensing.

@@

%\subsection{Observing Processes in the Upper Ocean}

% The processes occurring in the upper water-column and the photic zone
% (upper sunlit zone) are of vital importance for long-term sustainable
% management of the greater marine ecosystem, climate, and human health.
% With the onset of the warming oceans and increased anthropogenic
% pressure, it is critical to improve our understanding of the associated
% ecosystem response.
% owing to current observation practices.

Central to understanding the changes taking place in the upper
water-column is knowledge of the biophysical interaction driven by
an agglomeration of physical forcings %(wind, topography, bathymetry,
%tidal influences, and local geography, such as fresh water inflows) 
and incipient micro-biology driven by planktonic and coastal anthropogenic
input, such as pollution and agricultural runoff transported into the ocean by rivers and streams. These often result in a range of ecosystem-related phenomena such as blooms%, anoxic zones
and plumes. %, all of which have significant direct and indirect effects on society.

% On observation and why robotics
Traditional data collection to understand such phenomena at sea has typically been based on static
buoys, floats, or ship-based methods, with significant
logistical limitations that directly impact coverage and sampling
resolution. Modern methods using satellite remote-sensing provide
large-scale coverage but have limited resolution, are limited to
sensing the surface, and are impacted by cloud cover. Numerical ocean models similarly find it challenging to provide detail at
smaller scale \citep{Lermusiaux:2006}. 
The advent of mobile robotic platforms \citep{Bellingham07} has
resulted in significant contributions to environmental monitoring and
sampling. In particular, AUVs have
advanced the state of in-situ sampling, and made robotics an
integral part of ocean observation, filling parts of the undersampling gap (See Fig. \ref{fig:envir1}). %The previously listed challenges related to undersampling have also been driving ocean sensing practices towards more autonomous, networked, and distributed use of robotic assets (See Fig. \ref{fig:envir1}), where observation is guided by assimilation of recent measurements and statistical models.

@@

The AUVs considered here
operate at $1$-$3$ m/s in the upper water column. The in-water operation time capacity depends on survey speed, payload sensors and navigation. AUVs have limited computational capacity, and they are usually done on a single-board computer (SBC), like the Raspberry Pi or an multicore GPU NVIDIA Jetson TX1.

%By deploying AUVs together with other resources, one can augment information from synthetic ocean models, remote-sensing data, or fixed-location sensors on buoys. 
Surveys with AUVs are usually limited to observations
along fixed transects that are pre-programmed by the human operator. This occurs on a spatial grid, called a waypoint graph. A more effective approach is to instead use onboard algorithms to continuously evaluate, update, and refine future sampling locations (sense-plan-act cycle in Fig. \ref{fig:envir1}), making the information
gathering \emph{adaptive} \citep{das11b,fossum18b,fossuminformation}. In doing so, the space of sampling opportunities is still limited by the waypoint graph, but the AUV has flexibility to 
modify its path at each waypoint based on what has been measured onboard
\citep{py10,Rajan12,Rajan12b}.

The application that is of focus in this work is towards spatial characterization of a frontal system generated by river
plumes. Fig. \ref{fig:nidelven} shows the survey area (Nidelva, Norway) where cold freshwater enters
from the river, creating a strong gradient in both temperature and
salinity. Because of the local topography and the Coreolis force
%\citep{coriolis1835memoire} 
the cold fresh water tends to flow near
land to the east, %. Depending on the river discharge, tidal effects,
%wind, and temperature differences, this boundary often gets
%distorted. Initial knowledge about the location and evolution of these
but the locations of these features are highly uncertain, making deterministic planning challenging.

%River plumes belong to a class of ocean processes that are local and
%act on smaller \emph{sub-mesoscale} extent (from $5 m^2$ -- $10 km^2$) where
%lateral spatial variability tends to dominate. 
%At larger \emph{mesoscale} ($>50 km^2$), both three-dimensional space and time
%dynamics are important and can shift substantially, while in the case
%At this scale of river plumes one can further often limit scope to covering only lateral near surface spatial elements, and the case appears very amenable for AUVs usage. 
%However, the methodological framework can be extended to higher dimensional situations in a similar manner.

%\begin{figure}[!h] 
%\centering 
%\includegraphics[width=0.85\textwidth]{Figures/pictures/c-updated.pdf}
%\caption{Frontal patterns off of the Nidelva river, Trondheim, Norway.}
%\label{fig:nidelven}
%\end{figure}

%\begin{figure}[!h]
%\centering
 % \subfigure[River front - Columbia River, Astoria, Oregon, US.]{\includegraphics[width = 0.45\textwidth]{Figures/pictures/a.png}\label{fig:river1}}
 % \hfill
 %\subfigure[Tidal front - Korsfjorden, Norway.]{\includegraphics[width = 0.45\textwidth]{Figures/pictures/d.png}\label{fig:river2}}
 % \hfill
 % \subfigure[River plume - Rio de la Plata, Buenos Aires, Argentina.]{\includegraphics[width = 0.45\textwidth]{Figures/pictures/b.png}\label{fig:river3}}
 % \hfill
 %\subfigure[Frontal patterns - Nidelven, Trondheim, Norway.]{\includegraphics[width = 0.45\textwidth]{Figures/pictures/c.png}\label{fig:river4}}
 %\caption{Two examples of frontal features associated with
%   river-ocean interaction. % (\ref{fig:river1}) The river plume front of
   % the Columbia River meeting the Pacific Ocean.
%   (\ref{fig:river2})  Superimposed images (over 10 minutes) of a surface tidal front in
%   Korsfjorden, Norway. % (\ref{fig:river3}) The river plume of the Rio de
   % la Plata, taken by European Space Agency (ESA)'s Envisat platform and
   % MEdium Resolution Imaging Spectrometer (MERIS) sensor.
%   (\ref{fig:river4}) Aerial image of the mouth of the Nidelva river in
%   Trondheim, Norway where fresh water meets the salty fjord and creates
%   frontal patterns.} % Image courtesy: (\ref{fig:river1})
   % \cite{zamon2014marine}, (\ref{fig:river3}) ESA, CC BY-SA 3.0 IGO.}
% \label{fig:river_fronts}
%\end{figure}

%The use of \emph{robotic sampling} is motivated by the fundamental challenges of ocean observation, where a limited set of resources, a highly dynamic ocean, and the associated

%This challenge can be addressed by employing more elaborate and adaptive sampling strategies that can capitalize on both prior and current observations in order to locate and map these areas.

%\subsection{Autonomous Vehicles}

%To improve the state of sampling, modern tools and methods, including the use of autonomous platforms, oceanographic models and satellite remote sensing, needs to be combined with traditional data acquired by surface vessels or buoys (See Fig. \ref{fig:envir}). However, without adequate understanding of the theoretical underpinnings of how, when, and where to gather data, these tools and methods are insufficient in the vast and harsh oceans.

% retaining an advantageous strategy for information recovery, online during execution.
%The pressure on marine resources is growing and increased accuracy, resolution, and persistent monitoring of the oceans is crucial for long-term sustainable management. The impact of this research provides cost effective tools, techniques, and processes for doing ocean based measurements using robotic platforms. 
%adaptive design of experiments using robotic assets identifying and prioritizing relevant sampling locations on an information-theoretic basis. Spatial statistics naturally enters here through the ability to both model spatially correlated parameters and provide formal measures of uncertainty, on which this basis can be formed. 
%for oceanic sensing applications, where sensors are sparsely distributed and capitalizing on all available information is

%This imperfect understanding of causality coupled with uncertainty implies that we need to obtain more measurements and use statistical models and methods to determine what and how phenomena form, to increase the quality of predictive statements. To improve the state of sampling, modern tools and methods, including the use of autonomous platforms, oceanographic models and satellite remote sensing, augment the more traditional data acquired by surface vessels or buoys (See Fig. \ref{fig:envir}). However, without adequate understanding of the theoretical underpinnings of how, when, and where to gather data, these tools and methods are insufficient in the vast and harsh oceans.

%There are several oceanographic data sources that can be used together with statistical tools to improve data collection in ocean science. In the following section we briefly discuss buoy data, satellite data, ocean models data, and AUV data.

%Buoy data provide very accurate information of oceanographic variables, but in most situations they are local, giving information only at one (north, east) coordinate, possibly with opportunities for conducting measurements at different depths. Gliders and other surface vessels are a kind of floating buoys that drift and can measure variables at many locations, but still with limited spatial coverage.

%Satellite data are important for the mapping of oceanographic variables. They can also be indicative of variables such as temperature and salinity which we look at here, especially if the data are calibrated to for instance buoy data from the same spatial domain. However, the resolution of satellite data is relatively large, say $1 \times 1$ km $^2$ grid cells, and it only provides accurate information at the sea surface. Moreover, one cannot get useful satellite information on a cloudy day.

%What is often done in practice is to run ocean models based on the complex differential equations governing oceanographic phenomena, where also satellite data can be assimilated and used as input, and possibly playing with different forcing mechanisms to capture some of the uncertainty of the models. Initial conditions, and can be used t the goal is to guide the sampling towards regions these regions, where we can reduce the uncertainty in the ES. These models provide insight that can be understood from practitioners, but they tend to be biased, and must also be calibrated to buoy data in one way or the other.

%to provide an effective specification of regions of phenomenological interest resolving the boundaries and the position of phenomena, on which evaluation of future sampling can be constructed.

%; some of these phenomena are illustrated in Fig. \ref{fig:envir2}. Figure \ref{fig:river_fronts} illustrates the %phenomena of interest in this paper, which is river plumes and their interaction with the ocean.

%Feature driven sampling
%processes acting across a wide range of spatio-temporal scales makes prioritizing sampling efforts necessary. %Autonomous robotic platforms, can address these issues by providing the ability to focus sampling efforts to %high-interest regions (``hotspots").

% Determining paths for mobile robotic sensors in order to maximize the
% information gained about an environment is formally known as informative
% path planning. This, and related problems such as the orienteering
% problem \citep{Golden87}, has been studied in the context of graphs,
% where the potential measurement are assigned to nodes on which
% evaluation of different routes can be conducted.
%In this context, a fundamental result from \cite{nemhauser1978analysis} proves that a simple greedy algorithm (iteratively selecting the location which most increases the utility) can achieve a near-optimal solution if the problem can be shown to be \emph{submodular}\footnote{an intuitive diminishing returns property, where the informative value of adding sensors decrease with the number of sensors added.}. Building on this result, \cite{chekuri2005recursive} explored this in the setting of a graph using a recursive greedy algorithm, providing near-optimal solutions depending on the planning horizon and graph resolution.

@@

%Recent work on ESs and EPs connected to spatial statistics include
\cite{picheny2010,french2013spatio,bolin2015excursion,french2016credible}.
%Our focus is on ES as defined in Eq. (\ref{ES}) and the uncertainty
%reduction achieved by sampling as in Eq. (\ref{sur}). In this sense,
%our work is similar to \cite{bect2012}, \cite{chevalier2014fast}, and
%\cite{azzimonti2016quantifying} who describe analytical results for
%the IBV for univariate processes. A major contribution of our work is
%to derive closed-form results for this design criteria for situations
%where the underlying model is based on multivariate GPs.

@@

\end{document}

>>>>

\address{\\Trygve Olav Fossum \\Department of Marine Technology\\ Otto Nielsens veg. 10, 7491 Trondheim\\ Norway\\
\printead{e1}}
\address{Jo Eidsvik\\Department of Mathematical Sciences\\ Hogskoleringen 1, 7491 Trondheim\\ Norway\\ \printead{e2}}
\address{David Ginsbourger\\ Idiap Research Institute\\ Rue Marconi 19, 1920 Martigny\\ Switzerland\\
\printead{e3}}
\address{Kanna Rajan\\Underwater Systems and Technology Laboratory,
  Faculty of Engineering,\\ Rua Dr. Roberto Frias\\ University of Porto, Portugal\\

\printead{e4}}

@@

%Motivated by the challenges related to efficient allocation of sampling resources in environmental sensing, the combination of Excursion Probabilities and Gaussian process modeling is explored for autonomous robotic sampling of ocean features; enabling information driven measures  sampling efforts to high-interest regions. These regions are usually characterized by gradients of measurable environmental variables, e.g., temperature or salinity gradients, on which EPs subsequently can be used...Correlation among samples and multivariate requirements are typical in environmental studies.

\end{abstract}

@@

Monitoring the world's oceans has gained increased importance in 
light of the changing climate and increasing anthropogenic impact. A
central problem for understanding these factors is the lack of
representative data with sufficient resolution. Most of this
\emph{undersampling} can be attributed to the large spatio-temporal
variations on which ocean processes transpire, prompting the need for
effective means of sampling. By \emph{sampling}, we refer to the design of observational strategies in the spatial domain, where the use autonomous robotic platforms can be combined with statistical methods to pursue measurements with high scientific relevance. This combination, of statistical tools and robotic platforms, constitute a methodological basis for addressing the problem of efficient sampling of the ocean, which is at the core of this the paper.

There are several obstacles when doing sampling with autonomous platforms in the ocean. First, full numerical ocean models based on complex differential equations cannot provide accurate results for online operations because the onboard computers lack the computational power required. Hence statistical proxy models of the environment must be used. For our purpose, we rely on Gaussian process (GP) representations of the ocean variables of interest. Second, with undersampling issues one must usually plan for active learning during the operation, where a relevant criterion is used to extract the most valuable designs. The acquired in-situ information must then be assimilated into the statistical model onboard the platform, so that it can be used to re-compute the sampling criterion and in this way inform decisions on where to sample next. We use a criterion focusing on the prediction of excursion sets (ESs) for the multivariate spatial processes of interest. Third, the number of sampling designs is enormous, creating a trade-off between optimization and computability. One must often compromise in both development and practical operations, and we will rely on short-horizon optimization routines for sequential designs among nearest neighbors points in a spatial grid (waypoint graph).

The imperfect understanding of causality coupled with the underlying fundamental uncertainty of the ocean implies that we need to obtain more measurements and use statistical models and methods to determine what and how phenomena form, to increase the quality of predictive statements. With this in mind, we design AUV sampling strategies for reducing the spatially-integrated Bernoulli variance (IBV) in the ES indicator variables. Recent work on ESs and associated excursion probabilities (EPs) connected to spatial statistics include \cite{french2013spatio,bolin2015excursion,french2016credible}. Our focus is on uncertainty characterization and how this is influenced by sampling. In this sense, our work is similar to \cite{bect2012,chevalier2014fast,azzimonti2016quantifying} who describe analytical results for the IBV for univariate processes. A major contribution of our work is to derive closed-form results for this design criteria for situations where the underlying model is based on multivariate GPs. Another contribution is that we embed this approach in a sequential strategy to produce sampling algorithms, and apply this to a real-world scenario of autonomously sampling temperature and salinity variables using an autonomous underwater vehicle (AUV), with the intention of understanding the between the ocean and a river plume.

Motivating examples relevant for ESs of multivariate processes are abundant. In medicine, doctors do not rely solely on a single symptom but must see several combined effects before making a diagnosis. In our context of environmental sampling applications, the salinity and temperature excursions of a river plume can similarly help characterize the underlying oceanographic and bio-geochemical processes \cite{hopkins2013detection,Pinto2018}, where sampling should be aimed at mapping relevant parts of the domain to reduce uncertainty in the temperature and salinity ES. Unlike what has been done in previous plume exploration studies, we emphasize the links to spatial statistical modeling and targeted multivariate sampling criteria for sequential sampling.

@@

Central to understanding the changes taking place in the upper
water-column is knowledge of the biophysical interaction driven by
an agglomeration of physical forcings (e.g. wind, topography, bathymetry, tidal influences, etc.) and incipient micro-biology driven by planktonic and coastal anthropogenic input, such as pollution and agricultural runoff transported into the ocean by rivers and streams. These often result in a range of ecosystem-related phenomena such as blooms and plumes, with direct and indirect effects on society.

Traditional data collection to understand such phenomena at sea has typically been based on static buoys, floats, or ship-based methods, with significant logistical limitations that directly impact coverage and sampling resolution. Modern methods using satellite remote-sensing provide large-scale coverage but have limited resolution, are limited to sensing the surface, and are impacted by cloud cover. Numerical ocean models similarly find it challenging to provide detail at smaller scale \citep{Lermusiaux:2006}.  The advent of mobile robotic platforms \citep{Bellingham07} has resulted in significant contributions to environmental monitoring and sampling. In particular, AUVs have advanced the state of in-situ sampling, and made robotics an integral part of ocean observation, filling parts of the undersampling gap (See Fig. \ref{fig:envir1}).

@@

The AUVs considered here operate at $1$-$3$ m/s in the upper water column. The in-water operation time capacity depends on survey speed, payload sensors and navigation, see Fig. \ref{fig:lauv}. AUVs have limited computational capacity, and they are usually done on a single-board computer (SBC), like the Raspberry Pi or an multicore GPU NVIDIA Jetson TX1.

Surveys with AUVs are usually limited to observations along fixed transects that are pre-programmed by the human operator. This occurs on a spatial grid, called a waypoint graph. A more effective approach is to instead use onboard algorithms to continuously evaluate, update, and refine future sampling locations (sense-plan-act cycle in Fig. \ref{fig:envir1}), making the information gathering \emph{adaptive} \citep{das11b,Das2015,fossuminformation,fossum18b}. In doing so, the space of sampling opportunities is still limited by the waypoint graph, but the AUV has flexibility to  modify its path at each waypoint based on what has been measured onboard \citep{py10,Rajan12,Rajan12b}.

The application that is of focus in this work is towards spatial characterization of a frontal system generated by river plumes. Fig. \ref{fig:nidelven} shows the survey area (Nidelva, Norway) where cold freshwater enters from the river, creating a strong gradient in both temperature and salinity. Because of the local topography and the Coreolis force \citep{coriolis1835memoire} the cold fresh water tends to flow near
land to the east. Depending on the river discharge, tidal effects, wind, and temperature differences, this boundary often gets distorted. Initial knowledge about the location and evolution of these but the locations of these features are highly uncertain, making deterministic planning challenging.

@@

%Recent work on ESs and EPs connected to spatial statistics include
%\cite{picheny2010,french2013spatio,bolin2015excursion,french2016credible}.
%Our focus is on ES as defined in Eq. (\ref{ES}) and the uncertainty
%reduction achieved by sampling as in Eq. (\ref{sur}). In this sense,
%our work is similar to \cite{bect2012}, \cite{chevalier2014fast}, and
%\cite{azzimonti2016quantifying} who describe analytical results for
%the IBV for univariate processes. A major contribution of our work is
%to derive closed-form results for this design criteria for situations
%where the underlying model is based on multivariate GPs.

@@

% === Not used
% After reviews: - This is moved to Section 2.1.

%From an oceanographic perspective, interesting regions are usually directly tied to a distinct phenomena that is of scientific interest. Each phenomenon can in turn be characterized by a set of process specific conditions expressed through different measures of key environmental variables, such as temperature or salinity. One such measure is the gradient, that can be associated with a number of important processes, such as the vertical location of the thermocline and pycnocline, location of upwelling systems, vertical mixing, eddies, fronts, and currents \cite{sverdrup2006}, as well as distribution, growth, and accumulation of biological activity \cite{SatOceanSoci00, Ryan2014}. These gradients create boundaries separating the ocean into process specific regions which are of profound interest to both identify and map effectively. Quantification of gradient features is therefore a much needed competence in robotic sampling.
\end{document}
